Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-06-24 paper/paper.Rmd
move parts from blockCV buffering to general buffering intro

"Buffering" refers to the idea of removing observations between test and training set by drawing a circular buffer around one or more observations [@rest2014, @brenning2012, @blockCV].
By introducing more distance between observations, the influence of STAC should be reduced.
Buffering is mostly applied in the coordinate space due to the scattered appearance of observations in this space.
\pkg{mlr3spatiotempcv} supports two "buffering" implementations: (1) "Spatial buffering" from package \pkg{blockCV} and (2) "Leave-one-disc-out cross-validation" from \pkg{sperrorest}.

@@

<!--- (Subsection needs to be rewritten, see invisible comments. If this particular implementation is designed for binary classification problems, then a generic LOO with buffer is also needed, which should be introduced before this application-specific special case.) -->

<!--- P: By default the buffering implementation of blockCV behaves like a normal LOOCV.
It has additional features specific for SDM (presence-background data) but these are not active by default.
That is why I did not add `sperrorest::partition_loo()` because `blockCV::buffering()` does the same + has support for SDM. -->

<!-- A: This paragraph is a general description of the approach, not a description of unique characteristics of this implementation. General approach of buffering is (and should be) described in the introdoctory paragraph of the Buffering section.

The "buffering" method from the \pkg{blockCV} package is conceptually similar to a leave-one-out (LOO) cross-validation approach [@stone1974].
To reduce the similarity between training and test sets, a circular buffer around the observations representing the test set is drawn [@rest2014, @blockCV].
Within this buffer zone, all observations are removed from the training set, thereby reducing the similarity between training and test set.
-->

The implementation in the \pkg{blockCV} package comes with optional support for presence/absence data handling, as often used in the field of species distribution modeling (SDM) [@hijmans2013].
However, by default, the method considers all points for sampling the training and test sets.

In an ecological/species distribution modeling context, the default assumption of a binary response variable is to have *presence/absence* data.
This means all presence and absence values in the response variable are confirmed in their respective status.
<!--- FIXME: Why is this worth mentioning? What else could it be doing? LOO CV with buffer is a generic resampling method that can be applied to classification and regression problems. If this particular implementation only supports binary classification, then this should be mentioned explicitly. But mlr3 is not limited to binary classification problems, so why??? -->

<!--- P: See my comment above, it might be that this is a misunderstanding. If not (or anyhow), we might want to clarify this more than to prevent such misunderstandings. -->

The implementation also comes with support for *presence/background* data for binary responses.
In this scenario, only the presence values are considered validated. <!--- What is validated, in this context? Again, why overfit the code to such a very specific setting? -->

<!--- P: Validated here means: the existence of the presence observations is validated, the background observations are artificial. This is a common procedure in SDM.
The blockCV buffering integration comes with support for presence/background handling which is important for SDM.
I think this should be explained shortly then - which is what this section is about. -->

@@

<!--- again, why mention package name in title? -->

<!--- P: Because this is one method within the "buffering" category and so far my approach was to label these as "<method name> (<pkg>)".
Apparently this is sub-optimal from your point of view - how can we do better?
Also the visual appearance of level 4 headings in the PDF is sub-optimal - maybe we should do this someway different.
I've fixed the level 4 headers, they appear as proper formatted headers.
I also opened an issue on GH to add some introductory information to each method category. -->

@@

```{r custom-cv,  fig.cap="Environmental blocking using a single predictor.", out.width="50%", fig.pos="ht", eval=FALSE}
breaks <- quantile(task$data()$dem, seq(0, 1, length = 6))
zclass <- cut(task$data()$dem, breaks, include.lowest = TRUE)

>>>>

"Buffering" refers to the idea of removing observations between test and training set by drawing a circular buffer around one or more observations [@rest2014, @brenning2012, @blockCV].
Within this buffer zone, all observations are removed from the training set, thereby reducing the similarity between training and test set.

Buffering is mostly applied in the coordinate space due to the scattered appearance of observations in this space.
\pkg{mlr3spatiotempcv} supports two "buffering" implementations: (1) "Spatial buffering" from package \pkg{blockCV} and (2) "Leave-one-disc-out cross-validation" from \pkg{sperrorest}.

@@

The "buffering" method from the \pkg{blockCV} package is conceptually similar to a leave-one-out (LOO) cross-validation approach [@stone1974].
It comes with optional support for presence/absence data handling, as often used in the field of species distribution modeling (SDM) [@hijmans2013].
However, by default, the method considers all points for sampling the training and test sets.

In an ecological/species distribution modeling context, the default assumption of a binary response variable is to have *presence/absence* data.
This means all presence and absence values in the response variable are confirmed in their respective status.
The \pkg{blockCV} package implementation also comes with support for *presence/background* data for binary responses.

@@

@@

```{r custom-cv,  fig.cap="Environmental blocking using a single predictor.", out.width="50%", fig.pos="ht", eval=FALSE}
breaks = quantile(task$data()$dem, seq(0, 1, length = 6))
zclass = cut(task$data()$dem, breaks, include.lowest = TRUE)
